---
title: "Bycatch_apportionment"
author: "Jordan Watson"
date: "11/3/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
```

Some trips with only a single stat area have no salmon. 
Some trips with only a single stat area have salmon.
Some trips with multiple stat areas have no salmon.
Some trips with multiple stat areas have salmon. 

Catcher processors report their salmon at the haul level.
Shoreside delivers are reported at the trip level. 

Given the above observation of salmon and no-salmon in hauls and trips as described above, can we probabilistically apportion salmon within a given week to one stat area versus another when shoreside vessels deliver salmon after having fished in multiple stat areas?

```{r loadData}

data <- read_csv("rcode/Data/Genetic BSAI Salmon Bycatch.csv") %>% 
  rename_all(tolower) %>% 
  dplyr::select(-c(first_name,last_name,akfin_load_date)) %>% 
                  rename(malechumclip=`sum(male_chum_finclip)`,      
                         femalechumclip=`sum(female_chum_finclip)`,    
                         chumclip=`sum(total_chum_finclip)`,   
                         malechinclip=`sum(male_chinook_finclip)`,  
                         femalechinclip=`sum(female_chinook_finclip)`, 
                         chinclip=`sum(total_chinook_finclip)`,  
                         chin=`sum(number_chinook)`,         
                         chum=`sum(number_chum)`,            
                         coho=`sum(number_coho)`,           
                         pink=`sum(number_pink)`,            
                         sockeye=`sum(number_sockeye)`,     
                         unid=`sum(number_unidentified)`) %>% 
  mutate(salmon=chin+chum+coho+pink+sockeye+unid)


chindat <- data %>% 
  rename(stat=primary_adfg_stat_area_code,
         stat2=adfg_stat_area_codes) %>% 
         mutate(stat3=gsub(", ","",stat2),
                nstat=nchar(stat3)/6) %>% 
  dplyr::select(cruise,permit,season,haul_offload,week_number,processing_sector,catcher_vessel_adfg,
                stat,nstat,stat2,stat3,chin)
  
chindat %>% 
  ggplot(aes(factor(week_number),fill=processing_sector)) + 
  geom_bar(stat="identity") + 
  theme_bw()

unique(chindat$stat)
unique(chindat$stat2)
unique(chindat$processing_sector)

summary(chindat %>% filter(processing_sector=="S"))
```

```{r}
In a given week, how many of the hauls in a stat area were from trips where hauls occurred in only that stat area?
chindat %>% 
  filter(processing_sector=="S") %>% 
  ggplot(aes(week_number,fill=factor(nstat))) + 
  geom_bar(stat="count")
```

```{r SingleAreaTrips}
single <- chindat %>% 
  filter(nstat==1) %>% 
  dplyr::select(week_number,stats=stat,chin) %>% 
  group_by(week_number,stats) %>% 
  summarise(singlechin=sum(chin),
            singletrips=n(),
            singletripswchin=length(chin[chin>0])) %>% 
  mutate(stats=as.character(stats)) %>% 
  data.frame
```

```{r MultipleAreaTrips}
multi <- chindat %>% 
  filter(nstat>1) %>% 
  mutate(index=1:n())
```  

When we put these together, what does it look like? 

```{r}
newdat.a <- multi %>% 
  right_join(data.frame(index=sort(rep(multi$index,multi$nstat)),
                        stats=unlist(lapply(multi$index,function(x)strsplit(gsub(" ","",multi$stat2[x]), "[,]"))))) %>% 
  dplyr::select(-c(processing_sector,catcher_vessel_adfg,season,cruise,permit,stat3,nstat)) %>% 
  data.frame 

newdat.b <- newdat.a %>% 
  mutate(stats=as.character(stats)) %>% 
  group_by(week_number,stats) %>% 
  summarise(multitrips=n(),
            multichin=sum(chin)) 

#  There are a few trips that went to multiple areas and caught no salmon. We can use these to update the results of the single 
#  area data by increasing the number of trips that went to areas and caught no salmon. 

single <- bind_rows(newdat.b %>% 
            filter(multichin==0) %>% 
            rename(singletrips=multitrips,
                   singlechin=multichin) %>% 
            mutate(singletripswchin=0),
          single) %>% 
  group_by(week_number,stats) %>% 
  summarise(singlechin=sum(singlechin),
            singletrips=sum(singletrips),
            singletripswchin=sum(singletripswchin),
            propsingletripswbycatch=singletripswchin/singletrips)

#  Now remove these trips with no salmon from the multi-trip data
newdat.b <- newdat.b %>% 
  filter(multichin>0)

#  Join the two datasets. 
newdat <- newdat.b %>% 
  left_join(single) %>% 
  data.frame

#  To avoid accidentally tallying more zero salmon trips onto our datasets, remove newdat.b
rm(newdat.b)

```
                 
    week_number  stats multitrips multichin singlechin singletrips singletripswchin propsingletripswbycatch
1             4 625600          2        14        119          92               58              0.63043478
2             4 635600          2        14        170          54               45              0.83333333
3             4 645530          6       375        150           7                7              1.00000000
4             4 645600          8       389        135           5                5              1.00000000                 
                 
```{r}
newdat.a %>% filter(week_number==4 & grepl("625600",stat2))
single %>% filter(week_number==4 & stats%in%c("625600","635600"))
single2 %>% filter(week_number==4 & stats%in%c("625600","635600"))

multi %>% filter(week_number==4 & grepl("625600",stat2))

newdat.b %>% filter(multichin==0)
single %>% filter(week_number==30 & stats%in%c("675430","675500"))
single2 %>% filter(week_number==30 & stats%in%c("675430","675500"))

```
                    