---
title: "Bycatch_apportionment"
author: "Jordan Watson"
date: "11/3/2018"
output: html_document
---

Note: While this file refers to specific examples of bycatches in certain stat areas and during certain weeks,
I do not identify the year of the data anywhere. Thus, we retain the confidentiality in the data. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readxl)
```

Some trips with only a single stat area have no salmon. 
Some trips with only a single stat area have salmon.
Some trips with multiple stat areas have no salmon.
Some trips with multiple stat areas have salmon. 

Catcher processors report their salmon at the haul level.
Shoreside delivers are reported at the trip level. 

Given the above observation of salmon and no-salmon in hauls and trips as described above, can we probabilistically apportion salmon within a given week to one stat area versus another when shoreside vessels deliver salmon after having fished in multiple stat areas?

```{r loadData}

data <- read_csv("rcode/Data/Genetic BSAI Salmon Bycatch.csv") %>% 
  rename_all(tolower) %>% 
  dplyr::select(-c(first_name,last_name,akfin_load_date)) %>% 
                  rename(malechumclip=`sum(male_chum_finclip)`,      
                         femalechumclip=`sum(female_chum_finclip)`,    
                         chumclip=`sum(total_chum_finclip)`,   
                         malechinclip=`sum(male_chinook_finclip)`,  
                         femalechinclip=`sum(female_chinook_finclip)`, 
                         chinclip=`sum(total_chinook_finclip)`,  
                         chin=`sum(number_chinook)`,         
                         chum=`sum(number_chum)`,            
                         coho=`sum(number_coho)`,           
                         pink=`sum(number_pink)`,            
                         sockeye=`sum(number_sockeye)`,     
                         unid=`sum(number_unidentified)`) %>% 
  mutate(salmon=chin+chum+coho+pink+sockeye+unid)


chindat <- data %>% 
  rename(stat=primary_adfg_stat_area_code,
         stat2=adfg_stat_area_codes) %>% 
         mutate(stat3=gsub(", ","",stat2),
                nstat=nchar(stat3)/6) %>% 
  dplyr::select(cruise,permit,season,haul_offload,week_number,processing_sector,catcher_vessel_adfg,
                stat,nstat,stat2,stat3,chin)
  
chindat %>% 
  ggplot(aes(factor(week_number),fill=processing_sector)) + 
  geom_bar(stat="identity") + 
  theme_bw()

unique(chindat$stat)
unique(chindat$stat2)
unique(chindat$processing_sector)

summary(chindat %>% filter(processing_sector=="S"))
```

```{r}
In a given week, how many of the hauls in a stat area were from trips where hauls occurred in only that stat area?
chindat %>% 
  filter(processing_sector=="S") %>% 
  ggplot(aes(week_number,fill=factor(nstat))) + 
  geom_bar(stat="count")
```

```{r SingleAreaTrips}
single <- chindat %>% 
  filter(nstat==1) %>% 
  dplyr::select(week_number,stats=stat,chin) %>% 
  group_by(week_number,stats) %>% 
  summarise(singlechin=sum(chin),
            singletrips=n(),
            singletripswchin=length(chin[chin>0])) %>% 
  mutate(stats=as.character(stats)) %>% 
  data.frame
```

```{r MultipleAreaTrips}
multi <- chindat %>% 
  filter(nstat>1) %>% 
  mutate(index=1:n())
```  

When we put these together, what does it look like? 

```{r}
newdat.a <- multi %>% 
  right_join(data.frame(index=sort(rep(multi$index,multi$nstat)),
                        stats=unlist(lapply(multi$index,function(x)strsplit(gsub(" ","",multi$stat2[x]), "[,]"))))) %>% 
  dplyr::select(-c(processing_sector,catcher_vessel_adfg,season,cruise,permit,stat3,nstat)) %>% 
  data.frame 

newdat.b <- newdat.a %>% 
  mutate(stats=as.character(stats)) %>% 
  group_by(week_number,stats) %>% 
  summarise(multitrips=n(),
            multichin=sum(chin)) 

#  There are a few trips that went to multiple areas and caught no salmon. We can use these to update the results of the single 
#  area data by increasing the number of trips that went to areas and caught no salmon. 

single <- bind_rows(newdat.b %>% 
            filter(multichin==0) %>% 
            rename(singletrips=multitrips,
                   singlechin=multichin) %>% 
            mutate(singletripswchin=0),
          single) %>% 
  group_by(week_number,stats) %>% 
  summarise(singlechin=sum(singlechin),
            singletrips=sum(singletrips),
            singletripswchin=sum(singletripswchin),
            propsingletripswbycatch=singletripswchin/singletrips)

#  Now remove these trips with no salmon from the multi-trip data
newdat.b <- newdat.b %>% 
  filter(multichin>0)

#  Join the two datasets. 
newdat <- newdat.b %>% 
  left_join(single) %>% 
  data.frame

#  To avoid accidentally tallying more zero salmon trips onto our datasets, remove newdat.b
rm(newdat.b)

```

Using the above data, we can explore different relative weightings of stat areas. 

    week_number  stats multitrips multichin singlechin singletrips singletripswchin propsingletripswbycatch
1             4 625600          2        14        119          92               58              0.63043478
2             4 635600          2        14        170          54               45              0.83333333
3             4 645530          6       375        150           7                7              1.00000000
4             4 645600          8       389        135           5                5              1.00000000                 
                 
```{r}
newdat.a %>% filter(week_number==4 & grepl("625600",stat2))
single %>% filter(week_number==4 & stats%in%c("625600","635600"))
single2 %>% filter(week_number==4 & stats%in%c("625600","635600"))

multi %>% filter(week_number==4 & grepl("625600",stat2))

newdat.b %>% filter(multichin==0)
single %>% filter(week_number==30 & stats%in%c("675430","675500"))
single2 %>% filter(week_number==30 & stats%in%c("675430","675500"))

chindat %>% filter(nstat>1 & week_number==4 & stat%in%c("625600","635600","645600","645530")) %>% 
  dplyr::select(-c(season,stat3)) %>%
  mutate(stat=as.character(stat)) %>% 
  inner_join(newdat %>% 
               dplyr::select(propsingletripswbycatch,stat=stats,week_number))
```

#  Let's look at an example. We see that in week 4, there is a case were 7 chinook were caught in either stat area 625600 or 645600.
#  The primary stat area for the data is listed as 625600. However, when we look at newdat and filter by week 4 and stat areas 625600 and 645600,
#  we see that 5/5 trips (100%) caught Chinook in 645600 that week while only 58/92 (63%) of trips caught salmon in 625600.

```{r}
chindat %>% filter(nstat>1 & week_number==4 & stat%in%c("625600","635600","645600","645530")) %>% 
  dplyr::select(-c(season,stat3)) %>%
  mutate(stat=as.character(stat)) %>% 
  inner_join(newdat %>% 
               dplyr::select(propsingletripswbycatch,stat=stats,week_number))

newdat %>% filter(stats=="645600" & week_number==4)
newdat %>% filter(stats=="625600" & week_number==4)
```
                    
Let's look more explicitly at the primary stat area to which bycatch has been apportioned and the proportion of better observed trips with bycatch in those areas. 
```{r}                    
tempdat <- newdat.a %>% 
  dplyr::select(week_number,stats) %>% 
  inner_join(single) %>% 
  rename(stat=stats) %>% 
inner_join(chindat  %>% 
             dplyr::select(-c(season,stat3)) %>%
             mutate(stat=as.character(stat))) %>% 
  dplyr::select(-c(cruise,catcher_vessel_adfg)) %>% 
  filter(nstat>1 & chin>0) %>% 
  dplyr::select(permit,haul_offload,week_number,singlechin,singletrips,singletripswchin,chin,stat,stat2,propsingletripswbycatch)

tempdat %>% arrange(haul_offload) 

#  Need to spread the myval column so that there is a column for each of the different stat areas within a particular stat combo. 
#  This will yield up to 7 columns, I think, but in most cases there will just be two and it should make it easier to see different results.
newdat.a %>% 
  dplyr::select(week_number,stats,stat,stat2) %>% 
  inner_join(single) %>% 
  mutate(myval=paste0(round(propsingletripswbycatch,2)," (",stats,")"))





```                    